---
title: "accessory gland model"
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: "Ben Bolker and Lucas Eckert"
---

```{r setup, message=FALSE}
source("R/utils.R")
source("R/mcmc.R")
source("R/functions.R")
load_pkgs()
zmargin <- theme(panel.spacing = grid::unit(0, "lines"))
theme_set(theme_bw())
library(targets)
## tar_load(names=c(ag_compdata, ag_model0))
tar_load(everything())  ## incorporate all computed values
```

Approximate workflow:

```{r workflow, eval=FALSE}
## load phylo and trait data, replace NAs with ?, combine
get_ag_data("data/binaryTraitData.csv")
## set up constrained state matrix
ag_statemat1 <- with(ag_compdata, {
  ag_smdat <- corHMM::getStateMat4Dat(data)
  pars2equal <- list(c(7, 10, 20, 23), c(4, 11, 17, 24), c(2, 5, 15, 18), c(1, 8, 14, 21))
  equateStateMatPars(ag_smdat$rate.mat, pars2equal)
})
## run basic model and add state/parameter names
ag_model0 <- augment_model(
    corHMM(phy = ag_compdata$phy,
           data = ag_compdata$data,
           rate.cat = 1,
           rate.mat = ag_statemat1)
)
ag_mcmc0 <- 
  corhmm_mcmc(ag_model0,
              p_args=list(nllfun=make_nllfun(ag_model0)),
              n_cores = 8,
              n_chains = 8,
              ## these are per chain
              n_burnin = 4000,
              n_iter = 44000,
              n_thin = 10,
              seed = 101)
```

## some attempts at data viz

```{r mosaic, eval=FALSE, echo=FALSE}
## cool but useless
(ag_compdata$data
  %>% mutate(ag = paste0("ag:",ag),
             pc = paste0("pc:",pc),
             sc = paste0("sc:",sc))
  %>% 
  ggplot() + 
  geom_mosaic(aes(x = product(pc, sc), fill = ag)) +
  geom_mosaic_text(aes(x = product(pc, sc), fill = ag), as.label = TRUE) +
  scale_fill_discrete_qualitative()
)
```

less useless: 
```{r barchart}
(ag_compdata$data
  %>% mutate(across(c(pc, sc), factor, levels = c(0:1, "?")))
  %>% mutate(across(ag, factor, levels = 1:0)) ## reverse order to get ag=1 on top of bars
  %>% 
ggplot(aes(x=pc, fill = ag)) + geom_bar() +
  facet_wrap(~sc, labeller = label_both) +
scale_fill_discrete_qualitative()
)
```                         

##  priors

- all rates: from 1 (1 transition per tree) to 10*ntips (10 trans per species)
- `pc`: gain/loss ratio from 0.1 to 5
- `sc`: gain/loss ratio from 5 to 10
- `ag`: gain/loss ratio from 0.001 to 10


Use each of these as the basis for a +/- 3 sigma range (on the log scale)

```{r cifun, cache=TRUE}
t_list <- list(
    wald = tidy(ag_model0, conf.int = TRUE),
    ## profile = tidy(ag_model0, conf.int = TRUE,
    ## conf.method = "profile", profile = ag_profile0),
    mcmc = tidy(ag_mcmc0, conf.int = TRUE)
) 
all_ci <- (bind_rows(t_list, .id = "method")
  %>% rename(lwr = "conf.low", upr = "conf.high")
)
```

## Bayesian


```{r bayes_traceplot, fig.width=8, fig.height=8, cache=TRUE}
lattice::xyplot(ag_mcmc0, aspect="fill", layout=c(2,6))
```

```{r bayes_sum}
sort(effectiveSize(ag_mcmc0))
coda::gelman.diag(ag_mcmc0)
```

These diagnostics are OK but not great (ideally we would effective sample size up to 1000 and Gelman-Rubin ($\hat R$) statistic as low as 1.01 --- although older studies have used a higher cutoff). Probably fixable by "just" running the chain longer.

```{r bayes_pairs, fig.width=10, fig.height=10, echo=FALSE, eval=FALSE}
pairs(as.matrix(ag_mcmc0), gap = 0, pch = ".")
```

Not sure about the best way(s) to represent density: contour lines, filled contours ... change scale? Show highest posterior density contours?

```{r bayes_pairs2, fig.width=10, fig.height=10, cache=TRUE}
bp2 <- ggpairs(as.data.frame(lump.mcmc.list(ag_mcmc0)), progress=FALSE,
        lower=list(continuous=function(...) my_mcmc(..., show_prior=FALSE)),
        upper=list(continuous=function(...) my_mcmc(..., geom="density", show_prior=FALSE))) +
  zmargin
bp2_time <- system.time(print(bp2))
```

Contour levels are: 50%, 80% 90%, 95% (largest) highest posterior density regions.

### naming parameters

```{r image, message=FALSE}
image(ag_model0)
```

```{r parnames_manual}
m1L <- as.mcmc(ag_mcmc0)
m1L[,1] -> loss.sc
m1L[,2] -> loss.pc
m1L[,3] -> loss.ag_pc0_sc0
m1L[,4] -> gain.sc
m1L[,5] -> loss.ag_pc0_sc1
m1L[,6] -> gain.pc
m1L[,7] -> loss.ag_pc1_sc0
m1L[,8] -> loss.ag_pc1_sc1
m1L[,9] -> gain.ag_pc0_sc0
m1L[,10] -> gain.ag_pc0_sc1
m1L[,11] -> gain.ag_pc1_sc0
m1L[,12] -> gain.ag_pc1_sc1
```

Compare auto-names:

```{r parnames}
par_names(ag_model0)
```

### intermediate values

**FIXME**:  compare results here with contrast matrix results


```{r manual_contrasts}
#effect of pc on gaining ag
pc.gain.sc0 <- gain.ag_pc1_sc0 - gain.ag_pc0_sc0
pc.gain.sc1 <- gain.ag_pc1_sc1 - gain.ag_pc0_sc1
pc.gain <- (pc.gain.sc0+pc.gain.sc1)/2

#effect of pc on losing ag
pc.loss.sc0 <- loss.ag_pc1_sc0 - loss.ag_pc0_sc0
pc.loss.sc1 <- loss.ag_pc1_sc1 - loss.ag_pc0_sc1
pc.loss <- (pc.loss.sc0+pc.loss.sc1)/2

#effect of sc on gaining ag
sc.gain.pc0 <- gain.ag_pc0_sc1 - gain.ag_pc0_sc0
sc.gain.pc1 <- gain.ag_pc1_sc1 - gain.ag_pc1_sc0
sc.gain <- (sc.gain.pc0+sc.gain.pc1)/2

#effect of pc on losing ag
sc.loss.pc0 <- loss.ag_pc0_sc1 - loss.ag_pc0_sc0
sc.loss.pc1 <- loss.ag_pc1_sc1 - loss.ag_pc1_sc0
sc.loss <- (sc.loss.pc0+sc.loss.pc1)/2

#net effects on ag
pc.net <- pc.gain - pc.loss
sc.net <- sc.gain - sc.loss

df1<-data.frame(var="pc.gain",value=pc.gain)
df2<-data.frame(var="pc.loss",value=pc.loss)
df3<-data.frame(var="sc.gain",value=sc.gain)
df4<-data.frame(var="sc.loss",value=sc.loss)
df<-rbind(df1,df2,df3,df4)

print(ggplot(df, aes(x=var, y=var1))
      + geom_violin(fill="gray")
      + geom_hline(yintercept = 0, lty = 2)
      + coord_flip()
      )
```

interactions?

```{r int_hists}
int<-((gain.ag_pc1_sc1-loss.ag_pc1_sc1)-(gain.ag_pc0_sc1-loss.ag_pc0_sc1))-((gain.ag_pc1_sc0-loss.ag_pc1_sc0)-(gain.ag_pc0_sc0-loss.ag_pc0_sc0))
hist(int, breaks = 100)
```

```{r compute_contrasts, cache = TRUE}
cmat <- (read_csv("contr.csv", col_types = cols())
  ## arrange in same order as ag_mcmc1 columns!
  %>% mutate(across(parname, ~factor(., levels=colnames(ag_mcmc1))))
  %>% arrange(parname)
  %>% dplyr::select(-parname)  ## drop row name so we have a pure-numeric matrix
  %>% as.matrix()
)
ag_contr <- ag_mcmc1 %*% cmat
ag_contr_long <- (ag_contr
  %>% as_tibble()
  %>% pivot_longer(everything(), names_to = "contrast")
  %>% separate(contrast, into=c("contrast", "rate"))
)
```

**FIXME**: think more about net gain and how to characterize it.
Certainly should *not* be exponentiated, so leave it out here ...

```{r plot_contrasts, fig.height=4, fig.width=10}
ag_contr_gainloss <- ag_contr_long %>% filter(rate != "netgain")
ggplot(ag_contr_gainloss, aes(x = exp(value), y = rate)) +
  facet_wrap(~ contrast) +
  geom_violin(fill = "gray") +
  stat_summary(fun.data = "median_hilow", geom = "pointrange", colour="darkred") +
  geom_vline(xintercept = 1, lty = 2) +
  scale_x_log10()
```

```{r sum_tab, results="asis"}
bayes_pval <- function(x, ref=0) {
  x <- mean(x<ref)
  2 * min(x, 1-x)
}
(ag_contr_gainloss
  %>% group_by(contrast,rate)
  %>% summarise(median = median(value), lwr = quantile(value, 0.025), upr = quantile(value, 0.975), p = bayes_pval(value), .groups = "drop")
  %>% knitr::kable(digits = 3)
)
```

Graphing all CIs: bounds are set to the lower/upper bounds used for corHMM fit.

```{r ci_plot}
## order of parameters
levs <- c(outer(FUN = paste, sep = ".",
      c("loss", "gain"),
      c("sc", "pc",
        paste0("ag_", c(outer(paste0("pc", 0:1), paste0("sc", 0:1), paste, sep = "_"))))
      ))
corhmm_bounds <- log(c(lwr=0.1, upper = 100*ape::Ntip(ag_compdata$phy)))
100 * ape::Ntip(ag_compdata$phy)
all_ci <- all_ci %>% replace_na(list(upr = Inf, lwr = -Inf)) %>%
  mutate(across(term, factor, levels = levs))
## extend NA confidence interval to limit of graph
ggplot(
  all_ci,
  aes(term, exp(estimate), ymin = exp(lwr), ymax = exp(upr), colour = method)
) +
  geom_pointrange(position = position_dodge(width = 0.5)) +
  scale_colour_discrete_qualitative() +
  scale_y_log10(limits = exp(corhmm_bounds), oob = scales::squish) +
  coord_flip() +
  labs(y="rate", x = "")
```

```{r bayes_violin}
mcmc_long <- (ag_mcmc0
  %>% as.mcmc()
  %>% as_tibble()
  %>% pivot_longer(everything(), names_to = "term")
  %>% mutate(across(term, factor, levels = levs))
)

ggplot(mcmc_long, aes(x = exp(value), y = term)) +
  geom_violin(fill = "gray") +
  scale_x_log10() +
  stat_summary(fun.data = "median_hilow", geom = "pointrange", colour="darkred") +
  labs(x = "evolutionary rate", y = "")
```
