---
title: "accessory gland model: technical details"
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: "Ben Bolker and Lucas Eckert"
output:
  html_document:
    code_folding: hide
---

<!-- **fix me**: reimplement as [target markdown](https://books.ropensci.org/targets/markdown.html#markdown) ? -->

```{r setup, message=FALSE}
source("R/utils.R")
source("R/mcmc.R")
source("R/functions.R")
load_pkgs()
zmargin <- theme(panel.spacing = grid::unit(0, "lines"))
theme_set(theme_bw())
library(targets)
```

<!-- to suppress tidyverse messages, need callr_function = NULL *or* some other shenanigans -->

```{r visnetwork, message=FALSE, echo = FALSE, eval = FALSE}
tar_visnetwork(targets_only = TRUE)
```

##  priors

- all rates: from 1 (1 transition per tree) to 10*ntips (10 trans per species)
- `pc`: gain/loss ratio from 0.1 to 5
- `sc`: gain/loss ratio from 5 to 10
- `ag`: gain/loss ratio from 0.001 to 10


Use each of these as the basis for a +/- 3 sigma range (on the log scale)

```{r load}
tar_load(ag_model_pcsc)
tar_load(ag_mcmc0)
tar_load(all_ci)
tar_load(all_contr_ci)
```

## naming parameters

```{r image, message=FALSE}
plot(ag_model_pcsc)
```

## Contrasts matrix

This connects the parameters we estimate (e.g. `loss.ag_pc0_sc0`, loss rate of accessory glands when parental care and sperm competition are both absent) to the values we are interested in (e.g. `pc_loss`, effect of parental care on loss rate of accessory glands).

**fixme**: 

- reorder matrix for visualization?
- include weights by (imputed) state occupancy?


```{r plot_contrast_mat}
tar_load(contrast_mat)
## order nicely
order_fun <- function(x) {
    trait <- (stringr::str_extract(x,"\\.[[:alpha:]]{2}")
        |> stringr::str_remove("\\.")
        )
    tcode <- match(trait, c("pc", "sc", "ag"))
    ## loss/gain
    lossgain <- stringr::str_extract(x, "[[:alpha:]]{4}")
    lcode <- match(lossgain, c("loss", "gain"))
    rem <- stringr::str_remove(x,"[[:alpha:]]{4}\\.[[:alpha:]]{2}(_|$)")
    order(1000*tcode + 100*lcode + order(rem))
}
image_plot(t(contrast_mat[
    ## re-order; drop cols 1-4, which aren't involved in contrasts
    ##  (just gain/loss of pc, sc)
    order_fun(rownames(contrast_mat)),])[,-(1:4)],
    ylab = "AG contrast", xlab = "Rate parameter")
```

Distribution of states in the tree (computed from stochastic character mapping, histograms represent 100 simulations):

```{r states, warning=FALSE}
tar_load(states_df)
pivot_longer(states_df, everything(), names_to = "state") %>%
  ggplot(aes(x=value)) + geom_histogram(bins=25) + facet_wrap(~state) + zmargin
## sum columns then divide by total to get sum == 1
avg_states <- apply(states_df, 2, sum, na.rm = TRUE) / sum(states_df, na.rm = TRUE)
```

Need to figure out weights e.g. if we want the intercept for `loss.ag` it should be  something like:

```{r wts, eval=FALSE}
intercept_loss <- sum(occ.ag0_pc{i}_sc{j} * loss.ag_pc{i}_sc{j})
pc_loss <- (sum(occ.ag0_pc1_sc{i} * loss.ag_pc1_sc{i}) - sum(occ.ag0_pc0_sc{i} * loss.ag_pc0_sc{i})) /
  sum(occ.ag0_pc{i}_sc{j})
```

etc. ...

**FIXME**: think more about net gain and how to characterize it.
Certainly should *not* be exponentiated, so leave it out here ...

```{r contr, eval=FALSE}
## playing with contrasts definitions/scaling
dd <- expand.grid(f1=letters[1:2], f2 = LETTERS[1:2])
m <- model.matrix(~f1*f2, data=dd, contrasts = list(f1=contr.sum, f2=contr.sum))
t(solve(m))
## cols 2 and 3 have magnitude 1/4 but we want 1/2 (diff between treatments, not dev between
## ttt and mean)
```

```{r plot_contrasts, fig.height=4, fig.width=10}
tar_load(contr_long_ag_mcmc0)
tar_load(contr_long_ag_mcmc_tb)
ag_contr_gainloss <- purrr::map_dfr(list(fishphylo=contr_long_ag_mcmc0,
                                         treeblock=contr_long_ag_mcmc_tb),
                                    filter, rate != "netgain",
                                    .id = "phylo")

gg_sum <- ggplot(ag_contr_gainloss, aes(x = exp(value), y = rate)) +
  facet_wrap(~ contrast) +
  geom_violin(aes(fill = phylo), alpha=0.6) +
  stat_summary(fun.data = "median_hilow", geom = "pointrange", aes(group=phylo),
               ## width by trial and error; not sure what determines this?
               position = position_dodge(width=0.875),
               colour = "gray30") +
  geom_vline(xintercept = 1, lty = 2) +
  scale_x_log10() +
  zmargin +
  scale_fill_discrete_qualitative() +
  scale_colour_discrete_qualitative() +
  labs(x="expected transitions/proportional difference in rates")
print(gg_sum)
```

Note: values below are still on log scale ...

```{r sum_tab, results="asis"}
bayes_pval <- function(x, ref=0) {
  x <- mean(x<ref)
  2 * min(x, 1-x)
}
(ag_contr_gainloss
  %>% group_by(phylo,contrast,rate)
  %>% summarise(median = median(value), lwr = quantile(value, 0.025), upr = quantile(value, 0.975), p = bayes_pval(value), .groups = "drop")
  %>% knitr::kable(digits = 3)
)
```

**Conclusion**: treeblock and fishphylo results are nearly identical, treeblock results are *slightly* clearer.

Graphing all CIs: bounds are set to the lower/upper bounds used for corHMM fit. 
Vertical dashed lines show the mean and ± 3 SD of the priors.

```{r ci_plot, fig.width=12, fig.height = 6, warning = FALSE, message = FALSE}
tar_load(ag_compdata)
tar_load(all_ci)
## order of parameters
levs <- c(outer(FUN = paste, sep = ".",
      c("loss", "gain"),
      c("sc", "pc",
        paste0("ag_", c(outer(paste0("pc", 0:1), paste0("sc", 0:1), paste, sep = "_"))))
      ))
corhmm_bounds <- log(c(lwr=0.1, upper = 100*ape::Ntip(ag_compdata$phy)))
corhmm_prior <- log(c(lwr=1, upper = 10*ape::Ntip(ag_compdata$phy)))
corhmm_mid <- mean(corhmm_bounds)
fix_all_ci <- function(x, levels = NULL) {
    x <- (x
        ## replace NA values with Inf/-Inf so bars extend full range
        %>% replace_na(list(upr = Inf, lwr = -Inf))
        ## utility var for below-bounds estimates
        %>% mutate(lwr_bound = (estimate <= corhmm_bounds[["lwr"]]))
    )
    if (!is.null(levels)) {
        ## reorder levels as above
        x <- (x
            %>%  mutate(across(term, factor, levels = levels))
        )
    }
    return(x)
}

all_ci <- fix_all_ci(all_ci, levels = levs)

## extend NA confidence interval to limit of graph
gg_ci0 <- ggplot(
  all_ci,
  aes(term, exp(estimate), ymin = exp(lwr), ymax = exp(upr), colour = method)
) +
    geom_linerange(position = position_dodge(width = 0.5), key_glyph = "path") +
    geom_point(position = position_dodge(width = 0.5)) +
    colorspace::scale_colour_discrete_qualitative() +
    coord_flip() +
    labs(y="rate", x = "") +
    geom_hline(yintercept =exp(corhmm_mid), lty=2) +
    geom_hline(yintercept = exp(corhmm_prior), lty=2, col="gray") +
    theme(legend.position="bottom") +
    ## https://stackoverflow.com/questions/27130610/legend-on-bottom-two-rows-wrapped-in-ggplot2-in-r
    guides(colour = guide_legend(nrow = 2, byrow = TRUE, reverse = TRUE))
## would like to make arrows ("<") at lower bounds larger, but manipulating
##  size aesthetic would also mess up bars (I think). Some hack?
gg_ci1  <- (gg_ci0
    + scale_y_log10(limits = exp(corhmm_bounds), oob = scales::squish)
    + aes(shape = factor(lwr_bound))
    ## 60 = "<"
    + scale_shape_manual(values = c(16, 60), guide = "none")
)
gg_ci_nowald <- (gg_ci0 %+% dplyr::filter(all_ci, method != "model_pcsc")
    +  scale_y_log10()
)
cowplot::plot_grid(gg_ci1, gg_ci_nowald, nrow = 1)
```

- **prior_only**: 95% CIs implied by priors
- **model_pcsc**: full (12-parameter) frequentist (corHMM) model, unconstrained/unregularized
- **model_pcsc_prior**: corHMM model with priors (MAP estimates)
- **mcmc_0**: Bayesian MCMC fit with fully known phylogeny only (reduced taxa)
- **mcmc_tb**: Bayesian MCMC, tree block (all taxa, phylogenetic uncertainty)
- **mcmc_tb_nogainloss**: ditto, but without priors on gain/loss rates

## with contrasts

**FIXME**. Need to double-check the whole tidy/contrast pipeline. Are the results here consistent with
the main results (which are derived separately)?

```{r contrast_ci}
all_contr_ci <- fix_all_ci(all_contr_ci)
gg_ci1C  <- gg_ci1 %+% all_contr_ci
gg_ci_nowaldC <- gg_ci_nowald %+% all_contr_ci
cowplot::plot_grid(gg_ci1C, gg_ci_nowaldC, nrow = 1)
```

```{r mcmc_only, eval=FALSE}
## mcmc only
ggplot(
  all_ci %>% filter(method=="mcmc"),
  aes(y=term, x=exp(estimate), xmin = exp(lwr), xmax = exp(upr))) +
  geom_pointrange() +
  scale_x_log10()
```

```{r bayes_violin}
mcmc_long <- (ag_mcmc0
  %>% as.mcmc()
  %>% as_tibble()
  %>% pivot_longer(everything(), names_to = "term")
)  
```
