---
title: "Two Trait Model"
author: Lucas Eckert
output: html_document
---

The following are analyses of the correlation between parental care and accessory glands, done in 7 different ways. 

```{r, echo=F, include=F}
source("R/utils.R")
source("R/mcmc.R")
source("R/functions.R")
load_pkgs()
zmargin <- theme(panel.spacing = grid::unit(0, "lines"))
theme_set(theme_bw())
library(phytools)
```

```{r, echo=F}
allData <- read.csv("data/binaryTraitData.csv", header = TRUE)
traitData <- with(allData,
                  data.frame(species, ag, care, names=species))
fullPhy <- suppressWarnings(fishtree_phylogeny(traitData$species))
trimmedData <- comparative.data(fullPhy, traitData, names.col = names)
phy <- trimmedData$phy
data <- trimmedData$data
```

## 1. Pagel's Test

Pagel's 1994 test of correlated evolution. Creates independent and dependent models of evolution and compares them using a likelihood ratio test.
```{r, echo=F}
x<-data$ag
names(x)<-data$species
y<-data$care
names(y)<-data$species

phy2<-bind.tip(phy,tip.label="ROOT",edge.length=0,
    where=Ntip(phy)+1)

x<-as.factor(c(setNames(0,"ROOT"),
    setNames(as.character(x),names(x))))
y<-as.factor(c(setNames(0,"ROOT"),
    setNames(as.character(y),names(y))))

print(PagelsTest<-fitPagel(phy2, x, y, model = "ARD"))
```
P-value indicates correlation, rate estimates indicate a faster transition to accessory glands from a state of parental care compared to no care (about 10 fold).

## 2. Likelihood Ratio Test

This should be the same as Pagel's test. Independent and dependent models are created, this time using the same function we use to set up our Bayesian stuff later on.

Independent Model
```{r, echo=F}
LegendAndRateMat <- getStateMat4Dat(data)
RateMat <- LegendAndRateMat$rate.mat
pars2equal <- list(c(3,8), c(5,7), c(1,6), c(2,4))
indStateMat <- equateStateMatPars(RateMat, pars2equal)
indModel<- corHMM(phy = phy, data = data, rate.cat = 1, root.p = c(1,0,0,0), rate.mat = indStateMat)
print(indModel)
```

Dependent Model
```{r, echo=F}
depModel<- corHMM(phy = phy, data = data, rate.cat = 1, root.p = c(1,0,0,0))
print(depModel)
```

```{r, echo=F}
likRatio<- -2*(indModel$loglik-depModel$loglik)
p.val <- pchisq(likRatio, df = 4, lower.tail = F)
print(lr<-setNames(c(likRatio, p.val), c("likelihood ratio", "p-value")))
```
LR and p-value are the same as Pagel's test. Rates are very similar.

## 3. AIC Comparison

Same models as 1 and 2 except compared using AIC instead of log likelihoods.

AIC comparison for the models from Pagel's test, AIC values and weights
```{r, echo=F}
print(aic1<-setNames(c(PagelsTest$independent.AIC, PagelsTest$dependent.AIC),c("independant","dependent")))
aic.w(aic1)
```

AIC comparison for our corHMM models, values and weights
```{r, echo=F}
print(aic2<-setNames(c(indModel$AIC, depModel$AIC),c("independant","dependent")))
aic.w(aic2)
```
Strong evidence in favour of the dependent model in both cases.

## 4. BayesTraits (with BayesFactors)

Create the models in BayesTraits and then compare using BayesFactors, via the harmonic means of the log likelihoods. Harmonic means of ind, dep, and then the BF.

```{r, echo=F}
indBayesTraits<-read.csv("indBayesTraits.csv", header = T)
depBayesTraits<-read.csv("depBayesTraits.csv", header = T)

indLik<-1/mean(1/indBayesTraits$Lh)
depLik<-1/mean(1/depBayesTraits$Lh)
BF<-2*(depLik-indLik)
print(setNames(c(indLik, depLik, BF), c("independent log likelihood", "dependent log likelihood", "BayesFactor")))
```
BF of >10 indicates "very strong evidence" of correlation

## 5. BayesTraits (with contrasts)

Using the estimates from the dependent model from BayesTraits, contrasts on the effect of care on the gain, loss, and net-gain, of accessory glands

```{r, echo=F}
gain.ag.bt<-depBayesTraits$q24-depBayesTraits$q13
loss.ag.bt<-depBayesTraits$q42-depBayesTraits$q31
net.ag.bt<-gain.ag.bt-loss.ag.bt

df1<-data.frame(var="gain",value=gain.ag.bt)
df2<-data.frame(var="loss",value=loss.ag.bt)
df3<-data.frame(var="net",value=net.ag.bt)
df<-rbind(df1,df2,df3)

print(ggplot(df, aes(x=var, y=value))
      + geom_violin(fill="gray")
      + geom_hline(yintercept = 0, lty = 2)
      + coord_flip()
      )

print(ggplot(df, aes(x=var, y=value))
      + geom_violin(fill="gray")
      + geom_hline(yintercept = 0, lty = 2)
      + coord_flip()
      + ylim(c(-20,20))
      )
```

95% confidence intervals:
```{r, echo=F}
print(gain.bt.ci<-quantile(gain.ag.bt, c(0.025,0.975)))
print(loss.bt.ci<-quantile(loss.ag.bt, c(0.025,0.975)))
print(net.bt.ci<-quantile(net.ag.bt, c(0.025,0.975)))
```

Individual rates:
```{r, echo=F}
df4 <- (as.data.frame(depBayesTraits[,4:11])
  %>% pivot_longer(everything(), names_to="var", values_to = "value")
)

print(ggplot(df4, aes(x=var, y=value))
      + geom_violin(fill="gray")
      + geom_hline(yintercept = 0, lty = 2)
)

print(ggplot(df4, aes(x=var, y=value))
      + geom_violin(fill="gray")
      + geom_hline(yintercept = 0, lty = 2)
      + ylim(c(0,20))
)
```

## 6. BayesTraits (RJ)
The reversible jump method varies restrictions on the rates while attempting to find the best fitting model. To test our hypothesis, we can look at how often the rates towards the gain of ags are set to be the same and how often they are set to be different, depending on the state of parental care.
```{r, echo=F}
revJump<-read.csv("rjBayesTraits.csv", header = T)
rmat <- do.call("rbind",strsplit(revJump$Model.string, " "))
mean(rmat[,2] == rmat[,4])
```
The rates are the same about 75% of the time, not a lot of support for our hypothesis. 

## 7. Our corHMM Bayesian analysis
Same idea as the BayesTraits contrasts, except made using our method.
```{r, echo=F}
# data2<-na.omit(traitData[,1:3])
# treeblock<-fishtree_complete_phylogeny(data2$species)
# 
# depTreeBlock<- corHMM(phy = treeblock[[1]], data = data2, rate.cat = 1, root.p = c(1,0,0,0))
# 
# dep_treeblock_mcmc <-
#   corhmm_mcmc(depTreeBlock,
#               p_args=list(nllfun=make_nllfun(depTreeBlock), treeblock=treeblock),
#               n_cores = 3,
#               n_chains = 6,
#               ## these are per chain
#               n_burnin = 4000,
#               n_iter = 44000,
#               n_thin = 10,
#               seed = 101)
# saveRDS(dep_treeblock_mcmc, file = "depTreeBlockMCMC.rds")
```

```{r, echo=F, include=F}
dep_mcmc<-readRDS("depTreeBlockMCMC.rds")
lattice::xyplot(dep_mcmc, layout=c(2,4), aspect="fill",
                as.table = TRUE)
sort(effectiveSize(dep_mcmc))
coda::gelman.diag(dep_mcmc)
```

The contrasts for the effect of care on ag gain, loss, and netgain:
```{r, echo=F}
dep_mcmc[] <- lapply(dep_mcmc,
                       function(x) { 
                         colnames(x) <- c("q21", "q31", "q12", "q42",
                                          "q13", "q43", "q24", "q34")
                         return(x)
                       })
m1 <- as.mcmc(dep_mcmc)

gain<-m1[,7]-m1[,5]
loss<-m1[,4]-m1[,2]
net<-gain-loss

d1<-data.frame(var="gain",value=gain)
d2<-data.frame(var="loss",value=loss)
d3<-data.frame(var="net",value=net)
d<-rbind(d1,d2,d3)

print(ggplot(d, aes(x=var, y=var1))
      + geom_violin(fill="gray")
      + geom_hline(yintercept = 0, lty = 2)
      + coord_flip()
      )
```

The 95% confidence intervals:
```{r, echo=F}
print(gain.ci<-quantile(gain, c(0.025,0.975)))
print(loss.ci<-quantile(loss, c(0.025,0.975)))
print(net.ci<-quantile(net, c(0.025,0.975)))
```

The individual rates:
```{r, echo=F}
df0 <- (as.data.frame(m1)
  %>% pivot_longer(everything(), names_to="var", values_to = "logval")
  %>% mutate(value = exp(logval))
)

print(ggplot(df0, aes(x=var, y=value))
      + geom_violin(fill="gray")
      + geom_hline(yintercept = 0, lty = 2)
)

print(ggplot(df0, aes(x=var, y=value))
      + geom_violin(fill="gray")
      + geom_hline(yintercept = 0, lty = 2)
      + ylim(0,0.02)
)
```

